A Road Segment Prioritization Approach for Cycling Infrastructure

### 2023-09-22 README.Rmd
Add demographic data and surroundings

@@

Much research has explored the question of what gets people cycling.
Segregated cycling infrastructure --- road space that is allocated to cyclists only with physical separation to protect cyclists --- is associated with cycling uptake [@aldred2019impacts; @goodman2014new; @marques2015infrastructure].
Separation from motorized vehicles has found to be key [@winters2011motivators].
Revealed preference of cyclists shows that they are willing to deviate from the most efficient routes in order to commute on safer roads [@crane2017longitudinal].
However, such deviations are only considered if they do not considerably increase route circuitry; behavior studies have found that the probability of choosing a route decreases in proportion to its length relative to the shortest route [@broach2011bicycle; @winters2010far].
Another defining feature for cycling infrastructure is how well connected it is.
Cyclists prefer cohesive infrastructure, particularly when cycling on arterial roads with high levels of motorized traffic [@stinson2003commuter].
The lack of well-connected cycling infrastructure is one of the main obstacles to increasing cycling uptake [@caulfield2012determining].
While direct and cohesive cycling networks have been shown to positively impact cycling rates, density[^1] of the cycling network is also vital [@schoner2014missing].

@@

Other approaches have focused on characteristics of route segments as the basis of analysis to help prioritize investment in cycle infrastructure where it is most needed.
@grise2018if developed a multi-criteria approach, combined with estimated levels of usage, to highlight routes most in need of investment in Quebec City.
@wysling2022improve present a method that used open data from Paris to generate cycling suitability ratings based on 'bicycle suitability multipliers' associated with different types of road across the city, to support more effective interventions.

None of the aforementioned studies explore the equity impacts of different network prioritization approaches.
The following section highlights the importance of considering (often unstated) ethical assumptions underlying network optimisation and prioritization approaches of the type outlined above.

@@

Our work builds on these community finding approaches by proposing a similar greedy network expansion algorithm for cycle network expansion *within communities*.
We incorporate community finding methods for study area partitioning with weighted routing to avoid links that are stressful to cycle on.
In doing so, we propose an approach that ensures a better distribution investments across the study area while also accounting for motivators and deterrents to cycling.
We propose three sub-methods that address some of the limitations of previous studies.
These limitations include (a) an inherent bias when basing network design solely on existing cycling demand, (b) the proposal of routes that may not correspond to studies on cyclist preference and government policies, and (c) an insufficient consideration of the ethical principles underlying the analysis.
The current work addresses these limitations through the following organisation.
Section \@ref(calculating-potential-cycling-demand) focuses on calculating potential cycling demand.
Section \@ref(routing) focuses on routing the demand onto the road network while accounting for cyclist preferences and government priorities.
Section \@ref(road-segment-prioritization) outlines a method for partitioning the study area based on a community finding algorithm and routed cycling demand.
It then introduces the network expansion algorithms, and compares an approach grounded in 'egalitarianism' to one grounded in 'utilitarianism'. ithms, and compares an approach grounded in
'egalitarianism' to one grounded in 'utilitarianism'. As mentioned above, Manchester is used as a case study. 
An overview of the steps involved is shown in Figure \@ref(fig:methodflowchart).

@@

>>>>

```{r, include=FALSE}
library(tidyverse)
library(tmap)
knitr::opts_chunk$set(warning = FALSE)
```

@@

Much research has explored the question of what gets people cycling.
Segregated cycling infrastructure --- road space that is allocated to cyclists only with physical separation to protect cyclists --- is associated with cycling uptake [@aldred2019impacts; @goodman2014new; @marques2015infrastructure].
Separation from motorized vehicles has been found to be key [@winters2011motivators].
Revealed preference of cyclists shows that they are willing to deviate from the most efficient routes in order to commute on safer roads [@crane2017longitudinal].
However, such deviations are only considered if they do not considerably increase route circuitry; behavior studies have found that the probability of choosing a route decreases in proportion to its length relative to the shortest route [@broach2011bicycle; @winters2010far].
Another defining feature for cycling infrastructure is how well connected it is.
Cyclists prefer cohesive infrastructure, particularly when cycling on arterial roads with high levels of motorized traffic [@stinson2003commuter].
The lack of well-connected cycling infrastructure is one of the main obstacles to increasing cycling uptake [@caulfield2012determining].
While direct and cohesive cycling networks have been shown to positively impact cycling rates, density[^1] of the cycling network is also vital [@schoner2014missing].

@@

Other approaches have focused on characteristics of route segments as the basis of analysis to help prioritize investment in cycle infrastructure where it is most needed.
@grise2018if developed a multi-criteria approach, combined with estimated levels of usage, to highlight routes most in need of investment in Quebec City.
@wysling2022improve present a method that used open data from Paris to generate cycling suitability ratings, based on 'bicycle suitability multipliers' associated with different types of road across the city, to support more effective interventions.

With the exception of @grise2018if, the aforementioned studies do not explore the equity impacts of different network prioritization approaches.
The following section highlights the importance of considering (often unstated) ethical assumptions underlying network optimization and prioritization approaches of the type outlined above.

@@

Our work builds on these community finding approaches by proposing a similar greedy network expansion algorithm for cycle network expansion *within communities*.
We incorporate community finding methods for study area partitioning with weighted routing to avoid links that are stressful to cycle on.
In doing so, we propose an approach that ensures a better distribution investments across the study area while also accounting for motivators and deterrents to cycling.
We propose three sub-methods that address some of the limitations of previous studies.
These limitations include (a) an inherent bias when basing network design solely on existing cycling demand, (b) the proposal of routes that may not correspond to studies on cyclist preference and government policies, and (c) an insufficient consideration of the ethical principles underlying the analysis.
The current work addresses these limitations through the following organisation.
Section \@ref(calculating-potential-cycling-demand) focuses on calculating potential cycling demand.
Section \@ref(routing) focuses on routing the demand onto the road network while accounting for cyclist preferences and government priorities.
Section \@ref(road-segment-prioritization) outlines a method for partitioning the study area based on a community finding algorithm and routed cycling demand.
It then introduces the network expansion algorithms, and compares an approach grounded in 'egalitarianism' to one grounded in 'utilitarianism'. 
An overview of the steps involved is shown in Figure \@ref(fig:methodflowchart).

@@

## Case study area

Manchester is a city in the North West of England.
"Manchester" can be used to refer to multiple things:

- Manchester City Council, which is a local government district that includes the city centre and surrounding areas.
- The Greater Manchester region, which is a metropolitan county that includes the city of Manchester and 9 other local government districts.
- The city of Manchester, which is used by visitors to refer to the city centre and surrounding areas. This definition if formalised in the Major Towns and Cities (TCITY) dataset provided by the Office for National Statistics.

We use the TCITY definition of Manchester, which was "developed specifically for the production and analysis of statistics" and which is shown in Figure \@ref(fig:manchestermap).
Manchester is a cosmopolitan city with a large university population.
As shown in Figure \@ref(fig:manchestermap) (right), the high density of the central area allows high levels of cycling, although the surrounding areas have low levels of cycling.
Manchester and its surrounding have many deprived areas.

```{r manchestermap, echo=FALSE, fig.cap="Maps of Greater Manchester (top) and the case study area (City of Manchester)"}
greater_manchester = pct::get_pct_zones(region = "greater-manchester", geography = "msoa")
greater_manchester = greater_manchester |>
  mutate(pcycle = bicycle / all * 100)
# names(greater_manchester)
msoas = sf::read_sf("./data/Manchester/msoas_geometry.shp")
msoas_union = sf::st_union(msoas)
greater_manchester_lads = greater_manchester |>
  group_by(lad_name) |>
  summarise()
?tm_scale_bar
m1 = tm_shape(greater_manchester_lads) +
  tm_polygons(col = "lad_name", palette = "Set1", alpha = 0.2, title = "") +
  tm_layout(legend.position = c("left", "bottom")) +
  tm_shape(msoas_union) +
  tm_borders(col = "black", lwd = 2) +
  tm_scale_bar(breaks = (0:3)*5)
msoas_centroids = sf::st_centroid(msoas)
pct_zones_in_manchester = greater_manchester[msoas_centroids, ] 
m2 = tm_shape(greater_manchester, bb = msoas) +
  tm_fill("pcycle", palette = "Blues", title = "% cycling to work\n(2011 Census)") +
  tm_layout(legend.position = c("left", "bottom"), legend.bg.color = "white", legend.bg.alpha = 0.5) +
  tm_shape(msoas_union) +
  tm_borders(col = "black", lwd = 2) +
  tm_scale_bar(breaks = 0:5)
tmap_arrange(m1, m2, nrow = 1)
```
